---
title: "Data - Visualization"
author: "Stefan Glogger"
date: "21 August 2017"
output: pdf_document
---

```{r, include=FALSE, cache=T, results='hide'}
knitr::knit_child("20 Data Derivations.Rmd")
```

# Data Visualization

We visualize the data (stocks and sentix). For consistency, we first specify general parameters on how to display each index and the time periods. 

## Function

Put everything in one function to plot.

```{r}
colsEvalDates <- c("red", "darkred", "green", "orange", "yellow")
names(colsEvalDates) <- datesEvalNames
```

Rectangle for Date periods:
store as function to keep structure similar to above (and store at same Place in environment)

```{r}
plotData <- function(x, title = "Indices"){
    
    # lines with data
    geomLineDataDAX <- function(x){
        parse(text = paste0("geom_line(data = ", x, ", aes(x = Datum, y = DAX, colour = \"DAX\"))")) 
    }
    geomLineDataTEC <- function(x){
        parse(text = paste0("geom_line(data = ", x, ", aes(x = Datum, y = TEC, colour = \"TEC\"))"))
    }
    geomLineDataESX50 <- function(x){
        parse(text = paste0("geom_line(data = ", x, ", aes(x = Datum, y = ESX50, colour = \"ESX50\"))")) 
    }
    geomLineDataSP5 <- function(x){
        parse(text = paste0("geom_line(data = ", x, ", aes(x = Datum, y = SP5, colour = \"SP5\"))")) 
    }
    geomLineDataNASDAQ <- function(x){
        parse(text = paste0("geom_line(data = ", x, ", aes(x = Datum, y = NASDAQ, colour = \"NASDAQ\"))"))
    }
    geomLineDataNIKKEI <- function(x){
        parse(text = paste0("geom_line(data = ", x, ", aes(x = Datum, y = NIKKEI, colour = \"NIKKEI\"))"))
    }
    geomLineDataBUND <- function(x){
        parse(text = paste0("geom_line(data = ", x, ", aes(x = Datum, y = BUND, colour = \"BUND\"))")) 
    }
    
    # rectangle for date period
    geomRectDateBear <- function(){
        parse(text = "geom_rect(aes(xmin = min(datesEvalBear), xmax = max(datesEvalBear), ymin = -Inf, ymax = Inf), alpha = 0.2 , fill = \"red\")")}
    
    geomRectDateRecNBER <- function(){
      parse(text = "geom_rect(aes(xmin = min(datesEvalRecNBER), xmax = max(datesEvalRecNBER), ymin = -Inf, ymax = Inf), alpha = 0.2 , fill = \"darkred\")")}
    
    geomRectDateBull <- function(){
        parse(text = "geom_rect(aes(xmin = min(datesEvalBull), xmax = max(datesEvalBull), ymin = -Inf, ymax = Inf), alpha = 0.2 , fill = \"green\")")}
    
    geomRectDateLast <- function(){
        parse(text = "geom_rect(aes(xmin = min(datesEvalLast), xmax = max(datesEvalLast), ymin = -Inf, ymax = Inf), alpha = 0.2 , fill = \"orange\")")}
    
    geomRectDateTest <- function(){
        parse(text = "geom_rect(aes(xmin = min(datesTest), xmax = max(datesTest), ymin = -Inf, ymax = Inf), alpha = 0.2 , fill = \"blue\")")}
    
    
    ggplot() +
      eval(geomLineDataDAX(x)) +
      eval(geomLineDataTEC(x)) +
      eval(geomLineDataESX50(x)) +
      eval(geomLineDataNASDAQ(x)) +
      eval(geomLineDataNIKKEI(x)) +
      eval(geomLineDataBUND(x)) +
      eval(geomRectDateLast()) +
      eval(geomRectDateBear()) +
      eval(geomRectDateRecNBER())+
      eval(geomRectDateBull()) +
      eval(geomRectDateTest()) +
      labs(x = "Time", y = "Value") + 
      labs(title = title) +
      theme(plot.title = element_text(hjust = 0.5)) # align title in center
}

## if a special name is given, take it, otherwise take x (plot sentix by using same dataframe (adopted))
plotDataPDF <- function(x, xName = x){
    pdf(file.path(getwd(), "Plot", paste0(xName, ".pdf")), width = 10, height = 4)
    plot(plotData(x))
    dev.off()
}
```


TODO: environments in R, plug functions into environments to keep structure
<http://adv-r.had.co.nz/Environments.html>

```{r}
# ePlot <- new.env() # environment to store functions (doesn't work)

# ls.str(envir = ePlot)
```


probierer, funktioniert nicht (wollte alle linien auf einmal plotten)
```{r, run = FALSE}
# geomLineData <- function(x){
#     parse(text = paste0("eval(geomLineDataDAX(\"", x , "\")) + eval(geomLineDataTEC(\"", x , "\"))"))
# }
# 
# ggplot() +
#     eval(geomLineData("retPlot")) +
#     eval(geomRectDateLast) +
#     labs(x = "Time", y = "Value")
```


## Stocks

Start of with a value of 100 for each stock and then plot the evolvment of this stock.

### plot()

```{r}
retPlot <- matrix(100, nrow = nrow(stocks), ncol = ncol(stocks)-1)
retPlot[2:nrow(stocks), ] <- 1+ret # to multiply lateron, we have to add 1
retPlot <- apply(retPlot, 2, cumprod)
rownames(retPlot) <- stocks[,1]

xNames <- rownames(retPlot)
class(xNames) <- "Date"   # convert to date

cols <- rainbow(ncol(retPlot))
ylim <- c(min(retPlot), max(retPlot))
plot(xNames, retPlot[,1], type = "l", xlab = "Date", ylab = "Value", main = "Indices over time", 
     col = cols[1], ylim = ylim)
for(sentixGroup in 2:ncol(retPlot)){
    par(new=T)
    plot(xNames, retPlot[,sentixGroup], type = "l", col = cols[sentixGroup], axes = F, xlab="", ylab="", ylim = ylim)
}
legend("topleft", legend = colnames(stocks)[2:ncol(stocks)], col = cols, lty = 1)

rm(retPlot, xNames, ylim, sentixGroup)
```

### ggplot()

```{r}
library(ggplot2)
```

need data frame as input for ggplot

```{r, warning=FALSE}
retPlot <- matrix(100, nrow = nrow(stocks), ncol = ncol(stocks)-1)
retPlot[2:nrow(stocks), ] <- 1+ret # to multiply lateron, we have to add 1
retPlot <- apply(retPlot, 2, cumprod)

retPlot <- as.data.frame(retPlot)
colnames(retPlot) <- colnames(stocks)[2:ncol(stocks)]
retPlot$Datum <- stocks[,1]
class(retPlot$Datum) <- "Date"   # convert to date

cols <- rainbow(ncol(retPlot))
ylim <- c(min(retPlot[,1:(ncol(retPlot)-1)]), max(retPlot[,1:(ncol(retPlot)-1)]))


plotData("retPlot")
plotDataPDF("retPlot")

rm(retPlot, cols, ylim)
```



## Dispersion

Graphs can be found in "\\R-Research Project Statistics\\Plot Data".

```{r, warning=FALSE}
lateximport <- c(paste0("\\subsection{Herfindahl}"))

for(sentixGroup in names(sDisp)){
    title <- paste("sDisp", sentixGroup)
    
    sPlot <- sDisp[[sentixGroup]]
    plotDataPDF("sPlot", title)
    
    lateximport <- c(lateximport, paste0("\\includegraphics[width=\\textwidth]{",paste0(title, ".pdf"),"}\\linebreak"))
}

fileConnection <- file(file.path(getwd(), "Plot", paste0("0sentixDisp.txt")))
writeLines(lateximport, fileConnection)
close(fileConnection)

rm(sPlot, sentixGroup, lateximport, fileConnection)
```

And we provide summary statistics.

```{r}
lapply(sDisp, function(x) {base::summary(x[,-1], digits = 2)})
```


## Herfindahl

Graphs can be found in "\\R-Research Project Statistics\\Plot Data".

```{r, warning=FALSE}
lateximport <- c(paste0("\\subsection{Herfindahl}"))

for(sentixGroup in names(sHerf)){
    title <- paste("sHerf", sentixGroup)
    
    sPlot <- sHerf[[sentixGroup]]
    plotDataPDF("sPlot", title)
    
    lateximport <- c(lateximport, paste0("\\includegraphics[width=\\textwidth]{",paste0(title, ".pdf"),"}\\linebreak"))
}

fileConnection <- file(file.path(getwd(), "Plot", paste0("0sentixHerf.txt")))
writeLines(lateximport, fileConnection)
close(fileConnection)


rm(sPlot, sentixGroup, lateximport, fileConnection)
```

And we provide summary statistics.

```{r}
lapply(sHerf, function(x) {base::summary(x[,-1], digits = 2)})
```
