---
title: "Analysis"
author: "Stefan Glogger"
date: "August 2017"
output: pdf_document
---

```{r, include=FALSE, cache=T, results='hide'}
knitr::knit_child("30 Data Visualization.Rmd")
```

# Analysis

## Sentix to Returns

We first want to look on how dispersion affects returns. We hypothesize that (future) return is higher if dispersion is lower. Therefore, we look at the mean return of the *q* quantil of dispersion and the mean return of its *1-q* quantil. We do this for all the sentiments in comparison with all stocks. 

We depict this value. row: index, column: sentiment
We also depict the ranks (higher rank = higher value). Be careful as the absolute values of returns are different across indices (therefore ranking is not really justified) Comparing in each row is justified and higher value is good (should expecially be greater than 0).

Let *n* be the number of periods considered (n=1: just this period, n=2: this and next period) and let *m* be the time lapse (m=0: returns starting right now, m=1: returns starting 1 one period behind).

We specify the data (sDisp or sHerf) to allow more general access.

```{r}
q <- 0.1

compareSentRet <- function(n, m=0, dat){
    res <- matrix(NA, nrow = ncol(ret), ncol = length(dat))
    rownames(res) <- colnames(ret)
    colnames(res) <- names(dat)
    
    for(d in 1:length(dat)){
        for(s in 1:ncol(ret)){
            tmp <- data.frame(disp = dat[[d]][2:(nrow(dat[[d]])-n+1-m),s+1]) # 2 as returns start one date behind
            for(k in 1:(nrow(ret)-n+1-m)){
                tmp[k,"r"] <- prod(1+ret[(k+m):(k+m+n-1),s])-1
            }
            tmp <- tmp[order(tmp$disp),] # ascending by default
            res[s, d] <- round( mean(tmp[1:(q*nrow(tmp)),"r"]) - mean(tmp[((1-q)*nrow(tmp)):nrow(tmp), "r"]) , 3)
        }
    }
    return(res)
}
```



### actual dispersion to actual return, no lag

dispersion in connection with return of same period

So I1 seems to be able to predict returns, while P6 does not. 



```{r}
res <- compareSentRet(1, dat = sDisp)
res
matrix(rank(res), ncol = ncol(res), dimnames = list(rownames(res), colnames(res)))
```

### actual dispersion to actual return, lag of 1

dispersion in connection with return of same period


```{r}
res <- compareSentRet(1, 1, dat = sDisp)
res
matrix(rank(res), ncol = ncol(res), dimnames = list(rownames(res), colnames(res)))
```

### actual dispersion with future return (n=3), no lag

dispersion of one period with return over next *n* periods (this period up to n-1 period).

```{r}
res <- compareSentRet(3, dat = sDisp)
res
matrix(rank(res), ncol = ncol(res), dimnames = list(rownames(res), colnames(res)))

```

### actual dispersion with future return (n=6), no lag

dispersion of one period with return over next *n* periods (this period up to n-1 period).

```{r}
res <- compareSentRet(6, dat = sDisp)
res
matrix(rank(res), ncol = ncol(res), dimnames = list(rownames(res), colnames(res)))
```





### comparison overview general

m lag: 0, 2, 4
n periods considered: 1, 3, 6

generate an overview of dispersion with different periods considered in rows and different lags in columns. Each cell contains the results of the quantile difference for each stock.

function for comparison

```{r}
library(xtable)

compareOverview <- function(nVec, mVec, dat){
  res <- list()
  for(sentixGroup in names(dat)){
    rowVec <- integer()
    for(n in nVec){
      rowVec <- c(rowVec, n, rep("", ncol(ret))) # one empty line
    }
    res[[sentixGroup]] <- data.frame(Periods = rowVec)
  }
  
  # fill each column rowwise
  for(m in c(0,2,4)){
    # again reserve space first
    for(sentixGroup in names(dat)){
      res[[sentixGroup]] <- cbind(res[[sentixGroup]], setNames(data.frame(rep(c(colnames(ret), ""), length(nVec))), paste("Lag", m)))
      res[[sentixGroup]] <- cbind(res[[sentixGroup]], as.data.frame(matrix("", ncol=1, nrow=(length(nVec)*(ncol(ret)+1))), stringsAsFactors = F))
    }
    
    for(n in c(1,3,6)){
      tmp <- as.data.frame(compareSentRet(n, m, dat))
      
      # assign the values
      for(sentixGroup in names(dat)){
        res[[sentixGroup]][((ncol(ret)+1)*(which(n==nVec)-1)+1):((ncol(ret)+1)*which(n==nVec)-1), 
                           (1+(2*which(m==mVec)))] <- tmp[, sentixGroup]
      }    
      
    }
  }
  return(res)
}

compareOverviewTXT <- function(nVec, mVec, dat, fileName){
  res <- compareOverview(nVec, mVec, dat)
  
  lateximport <- c(paste0("\\subsection{", fileName, "}"))
  for(sentixGroup in names(sDisp)){
    lateximport <- c(lateximport, paste0("\\subsubsection{", sentixGroup, "}"))
    lateximport <- c(lateximport, print(xtable(res[[sentixGroup]]), floating = FALSE))
  }
  
  lateximport <- c(lateximport, "\\clearpage")
  
  fileConnection <- file(file.path(getwd(), "Plot", paste0("0Analysis", fileName,".txt")))
  writeLines(lateximport, fileConnection)
  close(fileConnection)
}
```

#### comparison overview dispersion

```{r}
# initialize
nVec <- c(1,3,6)
mVec <- c(0,2,4)

compareOverviewTXT(nVec, mVec, sDisp, "Dispersion")
```

#### comparison overview herfindahl

```{r}
# initialize
nVec <- c(1,3,6)
mVec <- c(0,2,4)

compareOverviewTXT(nVec, mVec, sHerf, "Herfindahl")
```


### cleanup
```{r}
rm(q, res, compareSentRet)
```