---
title: "Visualization"
author: "Stefan Glogger"
date: "August 2017"
output: pdf_document
---


```{r, include=FALSE, cache=T, results='hide'}
knitr::knit_child("50 Optimization.Rmd")
```

# Visualization

## Functions

### Evaluation of Varying Portfolio

We want to visualize the evolvement of a portfolio over each time window.

Be aware of the index shifting:
retPlot[j-1, i] take wealth of previous day
retOverTime[j-1,] take return of today (j is one step ahead)

Remove numbering of x-axis by *xaxt='n'*.

Generate *retPortSentixVarying*, the returns of portfolios with varying portfolio weights using sentix as third factor with optimal weights. It has the following structure:

time window -> dispersion (sentixGroup) -> return of Portfolio, sharpe ratio

x: portfolio weights
R: return of portfolio on each date
r: mean return of portfolio over whole time window
sd: standard deviation of return of portfolio over whole time window
sr: sharpe ratio (weekly)
anR: annualized return of portfolio over whole time window
anSd: annualized standard deviation
anSR: sharpe ratio (annual)
fweight: mean of goal function value
turnover: turnover of weights (how much of portfolio has to be changed) in percent
omega: Omega calculated by *PerformanceAnalytics*


Calculation of turnover:
NOT USEFUL (not comparable as portfolios evolve differently): we fix portfolio weights in t, hold these weights to t+1 (while portfolio raises to (1+ret)\*price_t), and may change weightsin t+1 -> amount changed is (change in weights) \* (value of index in t+1)
USEFUL: we calculate the percentage points of weights that change in each time step. Divide by 2 as a percentage point is taken from one part of the portfolio and given to another part (so counted twice) -> get amount of portfolio that changes

Use an adoption of *calcTestVar()*

```{r}
# install.packages("PerformanceAnalytics")
library(PerformanceAnalytics)
```

use VaR *Value at Risk* and ES *Estimated Shortfall* from *PerformanceAnalytics* with method "historical", as we have historical data.

```{r}
calcEvalVarClassic <- function(dat){
    res <- list()
    for(timeWindowName in names(dat)){
        timeWindow <- get(timeWindowName)
        retTimeWindow <- ret[timeWindow,]
        retTimeWindow <- retTimeWindow[-1,]
        colnames(retTimeWindow) <- colnames(ret)
        
        rf <- mean(retTimeWindow[,"BUND"])
        
        for(portfolioName in names(dat[[timeWindowName]])){
            R <- rowSums(dat[[timeWindowName]][[portfolioName]]$x * retTimeWindow) ## returns calculated old way
            
            turnover <- c(0, rowSums(abs(diff(dat[[timeWindowName]][[portfolioName]]$x)))/2) # start off with 0 to have same length
            VaR <- VaR(R, method = "historical")
            ES <- ES(R, method = "historical")
            
            r <- mean(R)
            sd <- sd(R)
            
            anR <- (1+r)^52-1
            anSd <- sqrt((sd^2)*52)
            
            omega <- Omega(R)
            cVaR <- CVaR(R)
            
            res[[timeWindowName]][[portfolioName]] <- list(x = dat[[timeWindowName]][[portfolioName]]$x, 
                                                           R = R, r = r, sd = sd, sr = r/sd,
                                                           anR = anR, anSd = anSd, anSR = anR/anSd, 
                                                           turnover = turnover, VaR = VaR, ES = ES,
                                                           omega = omega, cVaR = cVaR)
            
        }
    }
    return(res)
}
```

difference in function is that fweight is not there for classic portfolios
```{r}
calcEvalVarSentix <- function(dat){
    res <- calcEvalVarClassic(dat)
    for(timeWindowName in names(dat)){
        for(portfolioName in names(dat[[timeWindowName]])){
            fweight = mean(dat[[timeWindowName]][[portfolioName]]$obj)
            res[[timeWindowName]][[portfolioName]]$fweight <- fweight
        }
    }
    return(res)
}
```

### plot performance

We now optimize the plotting for ggplot(). (DOESN'T WORK)

Therefore our dataframe to plot should have the following structure:
date: Date
value: worth of Portfolio
portfolio: Portfolio (SentixGroup)

first in separate list, then in one dataframe

NOTE: returns occur one date later as stated here (in the data)


There has been an issue with *date*. It is getted as character and we need to transform it to integer and then back to date to store date as a numeric value (formatted as a date) and then used as x-axis

```{r}
plotPortfolio <- function(data, timeWindowName, bw=F){
    if(bw){
        datWork <- data[[timeWindowName]]
        timeWindow <- get(timeWindowName)
        
        retPlot <- data.frame(date = as.integer(as.Date(get(timeWindowName))))# date is read as character, we need double to plot for x-axis
        class(retPlot$date) <- "Date"
        
        for(s in names(datWork)){
            tmp <- cumprod(1+datWork[[s]]$R)
            if(s == "tanPort"){
                s <- "TP"
            }else if(s == "mVaPort"){
                s <- "MV"
            }else if(s == "rp"){
                s <- "RP"
            }else if(s == "PGMV"){
                s <- "GMV"
            }else if(s == "PMD"){
                s <- "MD"
            }else if(s == "ew"){
                s <- "EW"
            }
            retPlot[[s]] <- c(100, 100*tmp)
        }
        
        retPlotLong <<- gather(retPlot, portfolio, value, -date)
        retPlotLong$portfolio = factor(retPlotLong$portfolio, levels = c("P1", "P6", "I1", "I6", "G1", "G6", "TP", "MV", "RP", "GMV", "MD", "EW", "DAX", "TEC", "ESX50", "SP500", "NASDAQ", "NIKKEI", "BUND"))
        ggplot(retPlotLong, aes(x = date, y = value, col = portfolio, linetype = portfolio))+
            geom_line(lwd=1)+
            scale_linetype_manual(values = c(rep("solid", 2), rep("longdash", 2), rep("twodash", 2), "dotted", "dashed", "dotdash", "F1", "4C88C488", "12345678")) +
            scale_color_grey()+
            theme_bw() + 
            labs(x = "Time", y = "Value", colour = "Portfolio", linetype = "Portfolio")
  
    }else{
        datWork <- data[[timeWindowName]]
        timeWindow <- get(timeWindowName)
        
        retPlot <- data.frame(date = as.integer(as.Date(get(timeWindowName))))# date is read as character, we need double to plot for x-axis
        class(retPlot$date) <- "Date"
        
        for(s in names(datWork)){
            tmp <- cumprod(1+datWork[[s]]$R)
            if(s == "tanPort"){
                s <- "TP"
            }else if(s == "mVaPort"){
                s <- "MV"
            }else if(s == "rp"){
                s <- "RP"
            }else if(s == "PGMV"){
                s <- "GMV"
            }else if(s == "PMD"){
                s <- "MD"
            }else if(s == "ew"){
                s <- "EW"
            }
            retPlot[[s]] <- c(100, 100*tmp)
        }
        
        retPlotLong <<- gather(retPlot, portfolio, value, -date)
        retPlotLong$portfolio = factor(retPlotLong$portfolio, levels = c("P1", "P6", "I1", "I6", "G1", "G6", "TP", "MV", "RP", "GMV", "MD", "EW", "DAX", "TEC", "ESX50", "SP500", "NASDAQ", "NIKKEI", "BUND"))
        ggplot(retPlotLong, aes(x = date, y = value, col = portfolio))+
            geom_line(lwd=1)+
            theme_rgbw()+
            # theme_bw()+
            labs(x = "Time", y = "Value", colour = "Portfolio")
        
        
    #     # --------------------- OLD --------------------------------
    #     datWork <- data[[timeWindowName]]
    #     timeWindow <- get(timeWindowName)
    #     
    #     colBackground <- colsEvalDates[timeWindowName]
    #     
    #     retPlot <- data.frame(date = as.integer(as.Date(get(timeWindowName))))# date is read as character, we need double to plot for x-axis
    #     class(retPlot$date) <- "Date"
    #     
    #     for(s in names(datWork)){
    #         tmp <- cumprod(1+datWork[[s]]$R)
    #         retPlot[[s]] <- c(100, 100*tmp)
    #     }
    #     
    #     
    #     plotCommand <- paste0(text = "ggplot(retPlot, aes(x=date))+")
    #     for (i in 2:(ncol(retPlot)-1)){
    #         plotCommand <- paste0(plotCommand, "geom_line(aes(y=retPlot[,",i,"], color = colnames(retPlot)[",i,"])) + ")
    #     }
    #     plotCommand <- paste0(plotCommand, "geom_line(aes(y=retPlot[,",ncol(retPlot),"], color = colnames(retPlot)[",ncol(retPlot),"])) ")
    #     
    #     eval(parse(text = plotCommand))+  
    #         labs(title = paste("Time:", timeWindowName),
    #              y = "Value",
    #              x = "Date") +
    #         scale_color_discrete(name = "Index")+
    #         theme(panel.background = element_rect(fill = alpha(colBackground, 0.2)))
    }
}
```

```{r}
plotPortfolioComplete <- function(dat, fileName, bw=F){
    lateximport <- c(paste0("\\subsection{",fileName,"}"))
    
    for(d in datesEvalNames){
        plotPortfolio(dat, d)
        
        title <- paste0(fileName, "-", d, ".pdf")
        pdf(file.path(getwd(), "Plot", title), width = 10, height = 4)
        plot(plotPortfolio(dat, d, bw))
        dev.off()
        
        lateximport <- c(lateximport, paste0("\\includegraphics[width=\\textwidth]{",title,"}"))
    }
    
    fileConnection <- file(file.path(getwd(), "Plot", paste0("0",fileName,".txt")))
    writeLines(lateximport, fileConnection)
    close(fileConnection)
}
```


### change of weights

```{r}
plotWeightsLines <- function(datName, timeWindowName, s){
    dat <<- datName[[timeWindowName]][[s]]$x
    dat <<- as.data.frame(dat)
    dat$date <<- as.Date(rownames(dat))
    plotCommand <- paste0("ggplot(dat, aes(x=date)) +")
    
    colBackground <- colsEvalDates[timeWindowName]
	 
    for(i in 1:(ncol(dat)-2)){
        plotCommand <- paste0(plotCommand, "geom_line(aes(y=dat[,",i,"], color = colnames(dat)[", i, "])) +")
    }
    plotCommand <- paste0(plotCommand, "geom_line(aes(y=dat[,",ncol(dat)-1,"], color = colnames(dat)[", ncol(dat)-1, "]))")
 
    eval(parse(text = plotCommand))+
      ylim(0,1)+
      labs(title = paste("Time:", timeWindowName),
           subtitle = paste("Portfolio:", s),
           y = "Weight",
           x = "Date") +
      scale_color_discrete(name = "Index")+
      theme(panel.background = element_rect(fill = alpha(colBackground, 0.2)))
}
```

```{r}
plotWeightsLinesComplete <- function(dat, fileName){
    lateximport <- c(paste0("\\subsection{",fileName,"}"))
    
    for(timeWindowName in datesEvalNames){
        lateximport <- c(lateximport, paste0("\\subsubsection{", fileName, " - ", timeWindowName, "}"))
        
        for(portfolioName in names(dat[[timeWindowName]])){
            # plotWeightsLines(dat, d, s)
            
            title <- paste0(fileName, "-", timeWindowName,"-", portfolioName, ".pdf")
            pdf(file.path(getwd(), "Plot", title), width = 10, height = 4)
            plot(plotWeightsLines(dat, timeWindowName, portfolioName))
            dev.off()
            
            lateximport <- c(lateximport, paste0("\\includegraphics[width=\\textwidth]{",title,"}"))
        }
    }
    
    fileConnection <- file(file.path(getwd(), "Plot", paste0("0",fileName,".txt")))
    writeLines(lateximport, fileConnection)
    close(fileConnection)
}
```

### change of weights with turnover

to plot multiple plots on one plot <http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/>

```{r}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


```{r}
# Define 'dat' in global environment with "<<-" in order to circumvent the (new, i.e. after review) special plotting behaviour of ggplot2
# More details in https://stackoverflow.com/questions/10904124/global-and-local-variables-in-r
plotWeightsLinesWithTurnover <- function(datName, d, s){
    dat <<- datName[[d]][[s]]$x
    dat <<- as.data.frame(dat)
    dat$date <<- as.Date(rownames(dat))
    dat$turnover <<- datName[[d]][[s]]$turnover

    colBackground <- colsEvalDates[d]
        
    plotCommand <- paste0('ggplot(dat, aes(x=date)) +')
    for(i in colnames(dat)[1:(ncol(dat)-2)]){
      plotCommand <- paste0(plotCommand, 'geom_line(aes(y=dat[["',i,'"]], color = "', i, '")) +')
    }
    plotCommand <- substr(plotCommand, 1, nchar(plotCommand)-2)
    
    
    ## OLD code (from before review)
    # plotCommand <- paste0("ggplot(dat, aes(x=date)) +")
    # for(i in 1:(ncol(dat)-3)){
    #     plotCommand <- paste0(plotCommand, "geom_line(aes(y=dat[,",i,"], color = colnames(dat)[", i, "])) +")
    # }
    # plotCommand <- paste0(plotCommand, "geom_line(aes(y=dat[,",ncol(dat)-2,"], color = colnames(dat)[", ncol(dat)-2, "]))")
    
    
    
    p1 <- eval(parse(text = plotCommand))+
      ylim(0, 1)+
      labs(title = paste("Time:", d),
           subtitle = paste("Portfolio:", s),
           y = "Weight ",
           x = "Date",
           color = "Portfolios") +
      theme(panel.background = element_rect(fill = alpha(colBackground, 0.2)), legend.position = "bottom")
    p2 <- ggplot(dat, aes(x=date))+
      ylim(0,1)+
      geom_bar(aes(y=dat$turnover), stat = "identity", fill = "black")+
      labs(title = paste("Time:", d),
           subtitle = paste("Turnover:", s),
           y = "Turnover",
           x = "Date") +
      theme(panel.background = element_rect(fill = alpha(colBackground, 0.1)))

    multiplot(p1, p2)
}

 
# plotWeightsLinesWithTurnover(datName, d, s)

```

```{r}
plotWeightsLinesWithTurnoverComplete <- function(dat, fileName){
    lateximport <- c(paste0("\\subsection{",fileName,"}"))
    
    for(d in datesEvalNames){
        lateximport <- c(lateximport, paste0("\\subsubsection{", fileName, " - ", d, "}"))
        
        for(s in names(dat[[d]])){
            # plotWeightsLines(dat, d, s)
            
            title <- paste0(fileName, "-", d,"-", s, ".pdf")
            pdf(file.path(getwd(), "Plot", title), width = 10, height = 8)
            plotWeightsLinesWithTurnover(dat, d, s)
            dev.off()
            
            lateximport <- c(lateximport, paste0("\\includegraphics[width=\\textwidth]{",title,"}"))
        }
    }
    
    fileConnection <- file(file.path(getwd(), "Plot", paste0("0",fileName,".txt")))
    writeLines(lateximport, fileConnection)
    close(fileConnection)
}
```


### summary statistics

print the summary (in matrix to pass it on to LaTeX-Table lateron)

```{r}
summaryClassic <- function(datName, d, roundTo = 4){
    dat <- datName[[d]]
    
    mat <- matrix(NA, nrow = 6, ncol = length(dat))
    rownames(mat) <- c("Mean Return (an)", "Volatility (an)", "Sharpe Ratio (an)", "Omega", "VaR", "CVaR")
    colnames(mat) <- names(dat)
    
    for(sInd in 1:length(dat)){
        mat[1,sInd] <- round(dat[[sInd]]$anR, roundTo)
        mat[2,sInd] <- round(dat[[sInd]]$anSd, roundTo)
        mat[3,sInd] <- round(dat[[sInd]]$anSR, roundTo)
        mat[4,sInd] <- round(dat[[sInd]]$omega, roundTo)
        mat[5,sInd] <- round(dat[[sInd]]$VaR, roundTo)
        mat[6,sInd] <- round(dat[[sInd]]$cVaR, roundTo)
    }
    return(mat)
}
```

```{r}
# install.packages("xtable")
library(xtable)
```

"floating = FALSE" to eliminate table environment


```{r}
summaryClassicComplete <- function(dat, fileName, roundTo = 4){
    lateximport <- c(paste0("\\subsection{",fileName,"}"))
    
    for(d in datesEvalNames){
        lateximport <- c(lateximport, paste0("\\subsubsection{", d, "}"))
        lateximport <- c(lateximport, print(xtable(summaryClassic(dat, d, roundTo)), floating = FALSE))
        print(summaryClassic(dat, d, roundTo))
    }
    
    lateximport <- c(lateximport, "\\clearpage")
    fileConnection <- file(file.path(getwd(), "Plot", paste0("0",fileName,".txt")))
    writeLines(lateximport, fileConnection)
    close(fileConnection)
}
```


### whole analysis in one command

```{r}
wholeAnalysis <- function(dat, fileName){
    retDat <- calcEvalVarClassic(dat)
    
    # weights
    plotWeightsLinesWithTurnoverComplete(retDat, paste0("Weights-", fileName))
    
    # performance of portfolio
    plotPortfolioComplete(retDat, paste0("Performance-", fileName))
    
    # summary statistics
    summaryClassicComplete(retDat, paste0("Summary-", fileName))
}
```


## Classic Optimization

### Constant weights over time window

We want to visualize the evolvement of a portfolio over each time window.

Be aware of the index shifting:
retPlot[j-1, i] take wealth of previous day
retOverTime[j-1,] take return of today (j is one step ahead)

Remove numbering of x-axis by *xaxt='n'*.

```{r}
for(d in datesEvalNames){
    cols <- rainbow(length(xClassicConst[[d]]))
    retOverTime <- 1+ret[get(d),]
    retPlotDates <- get(d)
    retPlotDates <- c(datesAll[which(datesAll==min(retPlotDates))-1], retPlotDates)
    retPlot <- data.frame(Datum = retPlotDates)
    
    for(i in names(xClassicConst[[d]])){
        retPlot[1,i] <- 100
        for(j in 2:nrow(retPlot)){
            retPlot[j,i] <- retPlot[j-1,i]*crossprod(xClassicConst[[d]][[i]], retOverTime[j-1,])
        }
    }
    
    ylim = c(min(retPlot[,-1]), max(retPlot[,-1]))
    plot(retPlot[,2], type = "l", ylim = ylim, col = cols[1], main = d, xlab = "Date", ylab = "Value", xaxt='n')
    for(i in 3:ncol(retPlot)){
        par(new=T)
        plot(retPlot[,i], type = "l", ylim = ylim, axes = F, xlab = "", ylab = "", col = cols[i-1])
    }
    axis(1, at = c(0, 10, 20, 30, 40, 50), labels = retPlot[c(0, 10, 20, 30, 40, 50)+1,1])
    legend("bottomright", legend = names(xClassicConst[[d]]), col = cols, lty = 1)
    
    pdf(file.path(getwd(), "Plot", paste0("Performance-ClassicConst-", d, ".pdf")), width = 10, height = 8)
    plot(retPlot[,2], type = "l", ylim = ylim, col = cols[1], main = d, xlab = "Date", ylab = "Value", xaxt='n')
    for(i in 3:ncol(retPlot)){
        par(new=T)
        plot(retPlot[,i], type = "l", ylim = ylim, axes = F, xlab = "", ylab = "", col = cols[i-1])
    }
    axis(1, at = c(0, 10, 20, 30, 40, 50), labels = retPlot[c(0, 10, 20, 30, 40, 50)+1,1])
    legend("bottomright", legend = names(xClassicConst[[d]]), col = cols, lty = 1)
    dev.off()
}
```

### Varying of portfolio weights

```{r}
# install.packages("tidyr")
library(tidyr)
```


```{r}
wholeAnalysis(xClassicVar, "Classic")
```


### Varying of portfolio weights no risk free asset

```{r}
wholeAnalysis(xClassicVarNoRf, "Classic-No-Risk-Free")
```


## Sentix Optimization

```{r}
wholeAnalysis(xDispVarEval, "Sentix-Dispersion")
wholeAnalysis(xHerfVarEval, "Sentix-Herfindahl")
wholeAnalysis(xDispNoRFVarEval, "Sentix-Disperion-No-Risk-Free")
wholeAnalysis(xHerfNoRFVarEval, "Sentix-Herfindahl-No-Risk-Free")
```


## Sole Performance of Stock

We want to also plot the performance of the single indexes in the time window (100\% invested in one index and hold).

```{r}
xIndexEval <- list()
for(t in datesEvalNames){
  for(p in colnames(ret)){
    tmp <- matrix(0, ncol = ncol(ret), nrow = length(get(t))-1)
    colnames(tmp) <- colnames(ret)
    rownames(tmp) <- get(t)[1:(length(get(t))-1)]
        
    
    tmp[,which(colnames(ret) == p)] <- 1
    
    xIndexEval[[t]][[p]]$x <- tmp
  }
}
wholeAnalysis(xIndexEval, "Indices")
```


## All together

sentix with classic portfolio with varying weights

### classicWithRiskFree

```{r}
retPortClassicVarying <- calcEvalVarClassic(xClassicVar)
retPortSentixVarying <- calcEvalVarClassic(xDispVarEval)

retAllVarying <- retPortClassicVarying
for(timeWindowName in names(retAllVarying)){
    retAllVarying[[timeWindowName]] <- append(retAllVarying[[timeWindowName]], retPortSentixVarying[[timeWindowName]])
}



plotPortfolioComplete(retAllVarying, "Performance-All-classicWithRiskFree")
plotPortfolioComplete(retAllVarying, "Performance-All-classicWithRiskFree", bw=T)
summaryClassicComplete(retAllVarying, "SummaryAll-classicWithRiskFree")
```

### classicWithOutRiskFree

```{r}
retPortClassicVarying <- calcEvalVarClassic(xClassicVarNoRf)
retPortSentixVarying <- calcEvalVarClassic(xDispNoRFVarEval)

retAllVarying <- retPortClassicVarying
for(timeWindowName in names(retAllVarying)){
    retAllVarying[[timeWindowName]] <- append(retAllVarying[[timeWindowName]], retPortSentixVarying[[timeWindowName]])
}


plotPortfolioComplete(retAllVarying, "Performance-All-classicNoRiskFree")
summaryClassicComplete(retAllVarying, "SummaryAll-classicNoRiskFree")
```


```{r}
rm(retPortClassicVarying, retPortSentixVarying, retAllVarying)
```


## Sharpe Ratio comparison

We test if the sharpe ratio of the sentiment strategies is higher than the Sharpe Ratio of the classical portfolio optimization techniques. Use function *sharpeTesting* of package *PeerPerformance* and depict the p-values (should be that sentiment strategy does not have higher sharpe ratio than classical strategy).
row: Sentiment Strategy
column: Classical Strategy

```{r}
# install.packages("PeerPerformance")
library(PeerPerformance)
```


```{r}
printSharpeRatioComparison <- function(datSent, datClas, fileName){
    lateximport <- c(paste0("\\subsection{",fileName," - p-Values that sharpe of Row is not higher}"))
    
    sentCalc <- calcEvalVarClassic(datSent)
    clasCalc <- calcEvalVarClassic(datClas)
    
    for(timeWindowName in names(sentCalc)){
        lateximport <- c(lateximport, paste0("\\subsubsection{", timeWindowName, "}"))
        
        mat <- matrix(NA, nrow = length(names(sentCalc[[timeWindowName]])), 
                      ncol = length(names(clasCalc[[timeWindowName]])))
        rownames(mat) <- names(sentCalc[[timeWindowName]])
        colnames(mat) <- names(clasCalc[[timeWindowName]])
        
        for(portfolioName in names(sentCalc[[timeWindowName]])){
            rSent <- sentCalc[[timeWindowName]][[portfolioName]]$R
            
            for(portfolioName2 in names(clasCalc[[timeWindowName]])){
                rClas <- clasCalc[[timeWindowName]][[portfolioName2]]$R
                
                mat[which(names(sentCalc[[timeWindowName]]) == portfolioName),
                    which(names(clasCalc[[timeWindowName]]) == portfolioName2)] <- round(sharpeTesting(rSent, rClas)$pval, 4)
            }
        }
        lateximport <- c(lateximport, print(xtable(mat), floating = FALSE))
        print(mat)
    
    }
    
    lateximport <- c(lateximport, "\\clearpage")
    fileConnection <- file(file.path(getwd(), "Plot", paste0("0SharpeRatioComparison-",fileName,".txt")))
    writeLines(lateximport, fileConnection)
    close(fileConnection)
}
```

```{r}
printSharpeRatioComparison(xDispVarEval, xClassicVar, "Dispersion")
printSharpeRatioComparison(xHerfVarEval, xClassicVar, "Herfindahl")
```




## cleanup

```{r, eval=FALSE}
rm(calcEvalVarClassic, calcEvalVarSentix, plotPortfolio, plotPortfolioComplete, plotWeightsLines, plotWeightsLinesComplete, summaryClassic, summaryClassicComplete, wholeAnalysis)

detach("package:xtable", unload = T)
```

