---
title: "Further Calculations"
author: "Stefan Glogger"
date: "August 2017"
output: pdf_document
---


```{r, include=FALSE, cache=T, results='hide'}
knitr::knit_child("70 Visualization.Rmd")
```

```{r}
library(PerformanceAnalytics)
library(ggplot2)
library(xtable)
library(PeerPerformance)
```


# Further Calculations

## Parameters

We want to adopt for trading costs.

```{r}
costTrade = 0.0002
```

## Adopt portfolios for trading costs

We reuse most of the calcEvalVarClassiv function and just adopt slightly for the inclusion of trading costs (in the returns).

Turnover is calculated as the percentage of the portfolio that changes. This doesn't help us with the calculations. Therefore, we set things up new.

Calculate the portfolio returns new. Note: for portfolio optimization, the concrete value of underlying is irrelevant (just the weights are important). Therefore, we can accumulate the costs of transaction, to obtain a true value of the portfolio (less transaction costs) in the end.

buy/sell stocks at closing prices (actual day)



```{r}
calcEvalVarClassicTradCost <- function(dat){
    res <- list()
    for(timeWindowName in names(dat)){
        timeWindow <- get(timeWindowName)
        retTimeWindow <- ret[timeWindow,]
        colnames(retTimeWindow) <- colnames(ret)
        
        # stocks2 <- as.matrix(stocks[,2:ncol(stocks)])
        # rownames(stocks2) <- stocks[, 1]
        # stocksTimeWindow <- stocks2[timeWindow,]
        
        # rf <- mean(retTimeWindow[,"BUND"])
        
        for(portfolioName in names(dat[[timeWindowName]])){
            # CALCULATION OF R FIRST TRY
            # turnoverStocks <- abs(diff(dat[[timeWindowName]][[portfolioName]]$x))
            # turnoverStocks <- rbind(turnoverStocks, 0) # on last day, no turnover (keep portfolio)
            # 
            # costsTransact <- rowSums(stocksTimeWindow[-1,]*turnoverStocks*costTrade)
            # costsTransact <- c(0, costsTransact) # no costs for setting up portfolio
            # 
            # valuePort_Start <- rowSums(dat[[timeWindowName]][[portfolioName]]$x*stocksTimeWindow[-nrow(stocksTimeWindow),]) # start of day
            # valuePort_End <- rowSums(dat[[timeWindowName]][[portfolioName]]$x*stocksTimeWindow[-1,]) # end of day
            # 
            # R <- (valuePort_End)/(valuePort_Start)-1 ## returns without Trading Costs
            # R <- (valuePort_End - cumsum(costsTransact[-1]))/(valuePort_Start - cumsum(costsTransact[-length(costsTransact)]))-1 ## returns with Trading Costs at end of period
            # R <- (valuePort_End - c(0, cumsum(costsTransact[-c(1, length(costsTransact))])))/(valuePort_Start - c(0,0, cumsum(costsTransact[-c(1, length(costsTransact)-1, length(costsTransact))])))-1 ## returns with Trading Costs at start of period
            
            # R NEW VERSION; SHOULD BE CORRECT (COMPARE TO MAIL 8.11.)
            ## no transaction costs
            # R <- numeric(length(timeWindow)-1)
            # R <- rowSums(dat[[timeWindowName]][[portfolioName]]$x*(1+retTimeWindow[-1,]))/1-1
            ## with transaction costs (at start of period)
            turnoverStocks <- rowSums(abs(diff(dat[[timeWindowName]][[portfolioName]]$x)))
            R <- numeric(length(timeWindow)-1)
            R <- (rowSums(dat[[timeWindowName]][[portfolioName]]$x*(1+retTimeWindow[-1,])) - c(0, turnoverStocks)*costTrade)/1 - 1
            
            turnover <- c(0, rowSums(abs(diff(dat[[timeWindowName]][[portfolioName]]$x)))/2) # start off with 0 to have same length
            VaR <- VaR(R, method = "historical")
            ES <- ES(R, method = "historical")
            
            r <- mean(R)
            sd <- sd(R)
            
            anR <- (1+r)^52-1
            anSd <- sqrt((sd^2)*52)
            
            omega <- Omega(R)
            cVaR <- CVaR(R)
            
            res[[timeWindowName]][[portfolioName]] <- list(x = dat[[timeWindowName]][[portfolioName]]$x, 
                                                           R = R, r = r, sd = sd, sr = r/sd,
                                                           anR = anR, anSd = anSd, anSR = anR/anSd, 
                                                           turnover = turnover, VaR = VaR, ES = ES,
                                                           omega = omega, cVaR = cVaR)
            
        }
    }
    return(res)
}
```

difference in function is that fweight is not there for classic portfolios
```{r}
calcEvalVarSentixTradCost <- function(dat){
    res <- calcEvalVarClassicTradCost(dat)
    for(timeWindowName in names(dat)){
        for(portfolioName in names(dat[[timeWindowName]])){
            fweight = mean(dat[[timeWindowName]][[portfolioName]]$obj)
            res[[timeWindowName]][[portfolioName]]$fweight <- fweight
        }
    }
    return(res)
}
```

```{r}
wholeAnalysisTradCost <- function(dat, fileName){
    retDat <- calcEvalVarClassicTradCost(dat)
    
    # weights
    plotWeightsLinesWithTurnoverComplete(retDat, paste0("Weights-", fileName))
    
    # performance of portfolio
    plotPortfolioComplete(retDat, paste0("Performance-", fileName))
    
    # summary statistics
    summaryClassicComplete(retDat, paste0("Summary-", fileName))
}
```


### Varying of portfolio weights

```{r}
wholeAnalysisTradCost(xClassicVar, "Classic")
```


### Varying of portfolio weights no risk free asset

```{r}
wholeAnalysisTradCost(xClassicVarNoRf, "Classic-No-Risk-Free")
```


## Sentix Optimization

```{r}
wholeAnalysisTradCost(xDispVarEval, "Sentix-Dispersion")
wholeAnalysisTradCost(xHerfVarEval, "Sentix-Herfindahl")
wholeAnalysisTradCost(xDispNoRFVarEval, "Sentix-Disperion-No-Risk-Free")
wholeAnalysisTradCost(xHerfNoRFVarEval, "Sentix-Herfindahl-No-Risk-Free")
```


## Sole Performance of Stock

We want to also plot the performance of the single indexes in the time window (100\% invested in one index and hold).

```{r}
xIndexEval <- list()
for(t in datesEvalNames){
  for(p in colnames(ret)){
    tmp <- matrix(0, ncol = ncol(ret), nrow = length(get(t))-1)
    colnames(tmp) <- colnames(ret)
    rownames(tmp) <- get(t)[1:(length(get(t))-1)]
        
    
    tmp[,which(colnames(ret) == p)] <- 1
    
    xIndexEval[[t]][[p]]$x <- tmp
  }
}
wholeAnalysisTradCost(xIndexEval, "Indices")
```


## All together

sentix with classic portfolio with varying weights

### classicWithRiskFree

```{r}
retPortClassicVarying <- calcEvalVarClassicTradCost(xClassicVar)
retPortSentixVarying <- calcEvalVarClassicTradCost(xDispVarEval)

retAllVarying <- retPortClassicVarying
for(timeWindowName in names(retAllVarying)){
    retAllVarying[[timeWindowName]] <- append(retAllVarying[[timeWindowName]], retPortSentixVarying[[timeWindowName]])
}



plotPortfolioComplete(retAllVarying, "Performance-All-classicWithRiskFree")
summaryClassicComplete(retAllVarying, "SummaryAll-classicWithRiskFree")
```

### classicWithOutRiskFree

```{r}
retPortClassicVarying <- calcEvalVarClassicTradCost(xClassicVarNoRf)
retPortSentixVarying <- calcEvalVarClassicTradCost(xDispNoRFVarEval)

retAllVarying <- retPortClassicVarying
for(timeWindowName in names(retAllVarying)){
    retAllVarying[[timeWindowName]] <- append(retAllVarying[[timeWindowName]], retPortSentixVarying[[timeWindowName]])
}


plotPortfolioComplete(retAllVarying, "Performance-All-classicNoRiskFree")
summaryClassicComplete(retAllVarying, "SummaryAll-classicNoRiskFree")
```


```{r}
# rm(retPortClassicVarying, retPortSentixVarying, retAllVarying)
```


## Sharpe Ratio comparison

We test if the sharpe ratio of the sentiment strategies is higher than the Sharpe Ratio of the classical portfolio optimization techniques. Use function *sharpeTesting* of package *PeerPerformance* and depict the p-values (should be that sentiment strategy does not have higher sharpe ratio than classical strategy).
row: Sentiment Strategy
column: Classical Strategy

```{r}
# install.packages("PeerPerformance")
library(PeerPerformance)
```


```{r}
printSharpeRatioComparison <- function(datSent, datClas, fileName){
    lateximport <- c(paste0("\\subsection{",fileName," - p-Values that sharpe of Row is not higher}"))
    
    sentCalc <- calcEvalVarClassicTradCost(datSent)
    clasCalc <- calcEvalVarClassicTradCost(datClas)
    
    for(timeWindowName in names(sentCalc)){
        lateximport <- c(lateximport, paste0("\\subsubsection{", timeWindowName, "}"))
        
        mat <- matrix(NA, nrow = length(names(sentCalc[[timeWindowName]])), 
                      ncol = length(names(clasCalc[[timeWindowName]])))
        rownames(mat) <- names(sentCalc[[timeWindowName]])
        colnames(mat) <- names(clasCalc[[timeWindowName]])
        
        for(portfolioName in names(sentCalc[[timeWindowName]])){
            rSent <- sentCalc[[timeWindowName]][[portfolioName]]$R
            
            for(portfolioName2 in names(clasCalc[[timeWindowName]])){
                rClas <- clasCalc[[timeWindowName]][[portfolioName2]]$R
                
                mat[which(names(sentCalc[[timeWindowName]]) == portfolioName),
                    which(names(clasCalc[[timeWindowName]]) == portfolioName2)] <- round(sharpeTesting(rSent, rClas)$pval, 4)
            }
        }
        lateximport <- c(lateximport, print(xtable(mat, digits = 4), floating = FALSE))
        print(mat)
    
    }
    
    lateximport <- c(lateximport, "\\clearpage")
    fileConnection <- file(file.path(getwd(), "Plot", paste0("0SharpeRatioComparison-",fileName,".txt")))
    writeLines(lateximport, fileConnection)
    close(fileConnection)
}
```

```{r}
printSharpeRatioComparison(xDispVarEval, xClassicVar, "Dispersion")
printSharpeRatioComparison(xHerfVarEval, xClassicVar, "Herfindahl")
```


## Presenation of results

Generate table as Dominik did by hand previously.

```{r}
summaryClassic <- function(datName, d, roundTo = 4){
    dat <- datName[[d]]
    
    mat <- matrix(NA, nrow = 6, ncol = length(dat))
    rownames(mat) <- c("Mean Return (an)", "Volatility (an)", "Sharpe Ratio (an)", "Omega", "VaR", "CVaR")
    colnames(mat) <- names(dat)
    
    for(sInd in 1:length(dat)){
        mat[1,sInd] <- round(dat[[sInd]]$anR, roundTo)
        mat[2,sInd] <- round(dat[[sInd]]$anSd, roundTo)
        mat[3,sInd] <- round(dat[[sInd]]$anSR, roundTo)
        mat[4,sInd] <- round(dat[[sInd]]$omega, roundTo)
        mat[5,sInd] <- round(dat[[sInd]]$VaR, roundTo)
        mat[6,sInd] <- round(dat[[sInd]]$cVaR, roundTo)
    }
    return(mat)
}
```


```{r}
# Code-Schnipsel, wohl nicht vollständig
summaryTableDominik <- function(dat, fileName, roundTo = 4){
  lateximport <- c(paste0("\\subsection{",fileName," - summaryTableDominik}"))
        
    datesWeWant <- c("datesEvalBear", "datesEvalRecNBER", "datesEvalBull", "datesEvalAllAfterTest")
    
    tab <- matrix(0, nrow = 6*length(datesWeWant), ncol = 18)
    rownames(tab) <- rep(c("Ret. (an)", "Vol. (an)", "SR (an)", "Omega", "VaR", "CoVaR"), 4)
    
    for(d in datesWeWant){
        datInside <- dat[[d]]
        for(s in names(dat[[d]])){
            tab[(which(datesWeWant == d)-1)+1, which(names(dat[[d]]) == s)] = dat[[d]][[s]]$sr
        }
        tab <- c(lateximport, print(xtable(summaryClassic(dat, d, roundTo)), floating = FALSE))
        print(summaryClassic(dat, d, roundTo))
    }
    
    lateximport <- c(lateximport, "\\clearpage")
    fileConnection <- file(file.path(getwd(), "Plot", paste0("0",fileName,".txt")))
    writeLines(lateximport, fileConnection)
    close(fileConnection)
}
```


```{r}
retPortClassicVarying <- calcEvalVarClassicTradCost(xClassicVar)
retPortSentixVarying <- calcEvalVarClassicTradCost(xDispVarEval)
retPortHerfVarying <- calcEvalVarClassicTradCost(xHerfVarEval)

retAllVarying <- retPortClassicVarying
for(timeWindowName in names(retAllVarying)){
    retAllVarying[[timeWindowName]] <- append(retAllVarying[[timeWindowName]], retPortSentixVarying[[timeWindowName]])
}
for(timeWindowName in names(retAllVarying)){
    retAllVarying[[timeWindowName]] <- append(retAllVarying[[timeWindowName]], retPortHerfVarying[[timeWindowName]])
}


summaryTableDominik(retAllVarying, "SummaryDominik-classicNoRiskFree")
```

