---
title: "R-Code Visualizing of IR"
author: "Stefan Glogger"
date: "2 August 2017"
output: pdf_document
---



# Portfolio Analysis

```{r}
# library(PerformanceAnalytics)
```


## with sentiment

```{r}
evolve
```

```{r}
Eevolving <- list()
for(weight in names(E)){
    Eevolving[[weight]] <- lapply(E[[weight]], evolve)
}
```

somehow an lapply doesn't work, so I implement everything again

```{r}
Eanalysis <- list()
for(weight in names(Eevolving)){
    # Eanalysis[[weight]] <- lapply(Eevolving[[weight]], analysePortfolio) # TODO somehow not working
    l <- Eevolving[[weight]]
    RsClassic <- matrix(unlist(lapply(l, Return.calculate)), ncol = length(l)) # return from time to time
    RsTSClassic <- na.omit(xts(RsClassic, order.by = as.Date(datesEvalLast)))
    bench <- xts(rep(0, nrow(RsTSClassic)), order.by = as.Date(datesEvalLast)[-1]) # benchmark 0 everytime
    
    S1 <- as.matrix(table.AnnualizedReturns(RsTSClassic, Rf = bench)) # scale is chosen automatically to 'weekly'
    S2 <- -1*VaR(RsTSClassic) # 95% as default (outputs small negative value)
    
    ans <- rbind(S1, S2)
    colnames(ans) <- names(l)
    rownames(ans) <- c("Return (p.a.)", "StdDev. Risk (p.a.)", "Sharpe Ratio", "VaR 95% (p.a.)")
    
    Eanalysis[[weight]] <- round(ans,3)
}
```

```{r}
names(Eanalysis)[20]
Eanalysis[[20]]

names(Eanalysis)[100]
Eanalysis[[100]]

names(Eanalysis)[150]
Eanalysis[[150]]
```

### find best portfolio

we have to apply our goaling function again to come up with best portfolios
TODO do this directly when determining the portfolios

for all weights: go through all dates, add corresponding objective value to objective value and divide at the very end through the number of dates -> average objective value per weight and strategy

```{r}
wNames <- character(nrow(grid))
for(k in 1:length(wNames)){
    wNames[k] <- paste0(grid[k,1], "-", grid[k,2], "-", grid[k,3])
}

i <- sentixDataNames[1]
parse(text = paste0(i, "ObjValue <- data.frame(\"w1\" = grid[,1], \"w2\" = grid[,2], \"w3\" = grid[,3], \"h\" = 0)"))
for(i in sentixDataNamesReg){
    eval(parse(text = paste0(i, "ObjValue <- data.frame(\"w1\" = grid[,1], \"w2\" = grid[,2], \"w3\" = grid[,3], \"h\" = 0)")))
    eval(parse(text = paste0("rownames(", i, "ObjValue) <- wNames")))
}

for(d in datesEvalLast){
    d <- as.integer(d)
    dInd <- which(datesEvalLast==d)+1 # +1 to skip the first entry (start with equal weights)
    SSent <- cov(SentData[(which(rownames(SentData) == d)-sentLookback):
                              which(rownames(SentData) == d) - 1, ]) # -1 to just look in past
    rdat <- ret[unique(pmax(which(rownames(ret)<=d) - 1,1)),] # from beginning to one day in past
    muStock <- colMeans(rdat)
    SStock <- cov(rdat)
    
    for(i in sentixDataNamesReg){
        for(k in 1:nrow(grid)){
            w <- as.numeric(grid[k,])
            weight <- paste(w, collapse = "-")
            eval(parse(text = paste0(i, "ObjValue[weight,] <- ", i, "ObjValue[weight,] + hWeighted(E[[weight]][[i]][dInd,])")))   
        }
    }
}
```

```{r}
n <- length(datesEvalLast)
for(i in sentixDataNamesReg){
    eval(parse(text = paste0(i, "ObjValue[,4] <- ", i, "ObjValue[,4]/n")))
}
sentixG1dispRegObjValue
```

not useful as higher weight for sentix results in higher objective value

```{r}
for(i in sentixDataNamesReg){
    eval(parse(text = paste0(i, "ObjValue[,4] <- ", i, "ObjValue[,4]/", i, "ObjValue[,3]")))
}
sentixG1dispRegObjValue
```
objective value stays roughly the same after dividing through weight

### highest return

```{r}
load(file.path(folderData, "Optimization", "EDispersionMin_gloggest2017-08-16---11-35"))
Eanalysis <- E
```


```{r}
sentixReturns <- matrix(NA, nrow = length(Eanalysis), ncol = ncol(Eanalysis[[1]]))
rownames(sentixReturns) <- names(Eanalysis)
colnames(sentixReturns) <- colnames(Eanalysis[[1]])
for(i in 1:length(Eanalysis)){
    sentixReturns[i,] <- Eanalysis[[i]][1,]
}
sentixReturns
colMeans(sentixReturns)
```
```{r}
m1 <- apply(sentixReturns, 2, function(x) max(x))
m2 <- apply(sentixReturns, 2, function(x) which.max(x))
m3 <- rownames(sentixReturns)[apply(sentixReturns, 2, function(x) which.max(x))]

data.frame("maximum return" = m1, "index of max return" = m2, "weights for function" = m3)
```

```{r}
strategy <- which.max(m1)
weight <- m2[strategy]
sentPfMaxReturn <- Eevolving[[weight]][[strategy]]
```


### highest Sharpe Ratio

```{r}
sentixSharpe <- matrix(NA, nrow = length(Eanalysis), ncol = ncol(Eanalysis[[1]]))
rownames(sentixSharpe) <- names(Eanalysis)
colnames(sentixSharpe) <- colnames(Eanalysis[[1]])
for(i in 1:length(Eanalysis)){
    sentixSharpe[i,] <- Eanalysis[[i]][3,]
}
# sentixSharpe
colMeans(sentixSharpe)
```

```{r}
m1 <- apply(sentixSharpe, 2, function(x) max(x))
m2 <- apply(sentixSharpe, 2, function(x) which.max(x))
m3 <- rownames(sentixSharpe)[apply(sentixSharpe, 2, function(x) which.max(x))]

data.frame("maximum return" = m1, "index of max return" = m2, "weights for function" = m3)
```
```{r}
strategy <- which.max(m1)
weight <- m2[strategy]
sentPfMaxSharpe <- Eevolving[[weight]][[strategy]]
```





## without sentiment (classic)

Calculate Portfolio evolvement (start with 100, how does it evolve)

```{r}
EclassicEvolving <- lapply(Eclassic, evolve)
```


TODO: look at implementation of VaR (Pfaff multiplied value by -100)

```{r}
analysePortfolio
```


```{r}
analysePortfolio(EclassicEvolving)
```

## Comparison

```{r}
# install.packages("SharpeR") # installs many other packages as well
# library(SharpeR)
```


```{r}
Rsent <- Return.calculate(sentPfMaxSharpe)
Rclas <- Return.calculate(EclassicEvolving[["GMV"]])

table.AnnualizedReturns(Rsent)
table.AnnualizedReturns(Rclas)

sr_test(x = Rsent, y = Rclas, paired = TRUE,alternative="greater")
```



# Portfolio Visualization

```{r}
# library(grDevices) # for color
```


```{r}
# pdf(file = paste0(getwd(), "/Plot/optimal 08-08/", "optimal Portfolios all", ".pdf"), width = 12, height = 8)
cols <- topo.colors(4)
plot(EclassicEvolving[[1]], lwd = 1, ylab = "Index", xlab = "", 
     col = cols[1], main = "Comparison of Allocation Strategies", 
     ylim = c(min(unlist(lapply(EclassicEvolving, min)), sentPfMaxReturn, sentPfMaxSharpe),
              max(unlist(lapply(EclassicEvolving, max)), sentPfMaxReturn, sentPfMaxSharpe)))
for(i in 2:length(EclassicEvolving)){
    eval(parse(text = paste0("lines(EclassicEvolving[[", i, "]], col = cols[", i, "])")))
}
legend("topleft", legend = c(names(EclassicEvolving), "sMaxR", "sMaxS"), col = c(cols, "red", "orange"), lty = 1, lwd = 2)
abline(h = 100, col = "gray")

lines(sentPfMaxReturn, col = "red", lwd = 2)
lines(sentPfMaxSharpe, col = "orange", lwd = 2)
# dev.off()
```